---
title: "p8105_hw1_yw4657"
output: github_document
---

```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
```


# Problem 1

```{r load_data}
data("penguins", package = "palmerpenguins")
```

```{r check, eval=FALSE}
names(penguins)
nrow(penguins)
ncol(penguins)
mean(pull(penguins, flipper_length_mm), na.rm = TRUE)
```

The `penguins` dataset contains data on penguins from the Palmer
Archipelago. Important variables include `species`, `island`,
`bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`,
`sex`, and `year`. The dataset contains `r nrow(penguins)` observations
and `r ncol(penguins)` variables. The mean flipper length is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), 2)` mm.

```{r ggplot, warning=FALSE}
plot1 <- ggplot(penguins, aes(x = bill_length_mm, 
                           y = flipper_length_mm, 
                           color = species)) + geom_point()

plot1
```

```{r ggsave, warning=FALSE}
ggsave("penguins_scatterplot.png", plot = plot1)
```

# Problem 2

```{r mean}
set.seed(123)

sample = rnorm(10)

logical_var = sample > 0

character_var = c("A", "B", "C", "D", "E",
                  "F", "G", "H", "I", "J")

factor_var = factor(c("low", "medium", "high", "low", "medium",
                      "high", "low", "medium", "high", "low"))

df = data.frame(
  sample = sample,
  logical_var = logical_var,
  character_var = character_var,
  factor_var = factor_var
)
```

```{r}
mean(pull(df, sample))
```

```{r}
mean(pull(df, logical_var))
```

```{r, warning=FALSE}
mean(pull(df, character_var))
```

```{r, warning=FALSE}
mean(pull(df, factor_var))
```

The mean can be calculated for the numeric and logical variables,
but not for the character or factor variables. The logical variable
can be averaged because R treats `TRUE` as 1 and `FALSE` as 0.
Character values cannot be averaged because they are not numeric.
Although factors have internal numeric codes, a factor itself is not
a numeric variable, so its mean cannot be calculated directly.

```{r coercion, eval=FALSE}
as.numeric(pull(df, logical_var))
as.numeric(pull(df, character_var))
as.numeric(pull(df, factor_var))
```

Converting the logical variable to numeric produces 1s and 0s,
which explains why its mean can be calculated. Converting the
character variable produces `NA` values because the character
values are not numeric. Converting the factor produces its internal
numeric level codes rather than the original factor labels. These
results help explain why the numeric and logical variables can be
averaged directly, whereas the character and factor variables cannot.