---
title: "p8105_hw1_yw4657"
output: github_document
---

```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
```


# Problem 1

```{r}
data("penguins", package = "palmerpenguins")
```

```{r, eval=FALSE}
names(penguins)
nrow(penguins)
ncol(penguins)
mean(pull(penguins, flipper_length_mm), na.rm = TRUE)
```

The `penguins` dataset contains data on penguins from the Palmer
Archipelago. Important variables include `species`, `island`,
`bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`,
`sex`, and `year`. The dataset contains `r nrow(penguins)` observations
and `r ncol(penguins)` variables. The mean flipper length is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), 2)` mm.

```{r, warning=FALSE}
plot1 <- ggplot(penguins, aes(x = bill_length_mm, 
                           y = flipper_length_mm, 
                           color = species)) + geom_point()

plot1
```

```{r, warning=FALSE}
ggsave("penguins_scatterplot.png", plot = plot1)
```

# Problem 2

```{r}
set.seed(123)

sample = rnorm(10)

logical_var = sample > 0

character_var = c("A", "B", "C", "D", "E",
                  "F", "G", "H", "I", "J")

factor_var = factor(c("low", "medium", "high", "low", "medium",
                      "high", "low", "medium", "high", "low"))
```

```{r}
df = data.frame(
  sample = sample,
  logical_var = logical_var,
  character_var = character_var,
  factor_var = factor_var
)
```

```{r}
mean(pull(df, sample))
```

```{r}
mean(pull(df, logical_var))
```

```{r, warning=FALSE}
mean(pull(df, character_var))
```

```{r, warning=FALSE}
mean(pull(df, factor_var))
```

The mean can be calculated for the numeric and logical variables,
but not for the character or factor variables. The logical variable
can be averaged because R treats `TRUE` as 1 and `FALSE` as 0.
Character values cannot be averaged because they are not numeric.
Although factors have internal numeric codes, a factor itself is not
a numeric variable, so its mean cannot be calculated directly.

